---
title: Classwork 4 - Example Answers
subtitle: ggplot Basics
date: 2023-03-08
execute: 
  echo: true
  warning: false
  message: false
  
from: markdown+emoji
---

```{r}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Question 1. First Steps

Load the following data.frame, `mpg`.
```{r}
library(tidyverse)
mpg <- ggplot2::mpg
```


```{r}
datatable(mpg)
```


```{r}
rmarkdown::paged_table(mpg)
```


<br>

```{r}
skim(mpg)
```



## Q1a. 
Run `ggplot(data = mpg)`. What do you see?
```{r}
ggplot(data = mpg)
```

<br>


## Q1b.
Write a ggplot code to make a scatterplot of `hwy` vs. `cyl`.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```


- The above visualization is not great.

- When we describe the relationship between continuous and categorical variables, it is better to go with the visualization about how the distribution of a continuous variable varies by a categorical variable.

### Histogram - distribution of a continous variable

```{r}
p <- ggplot(mpg, aes(x = hwy))

p + geom_histogram(aes(fill = factor(cyl)), alpha = .3)
```

```{r}
p <- ggplot(mpg, aes(x = hwy))

p + geom_freqpoly(aes(color = factor(cyl)), alpha = .5)
```

```{r}
p <- ggplot(mpg, aes(x = hwy))

p + geom_density(aes(fill = factor(cyl)), alpha = .5) +
  scale_x_continuous(breaks = seq(5,40,5)) +
  scale_fill_manual(values = c("#ff6600", "#00ccff", "#556b2f", "#66023c" )) +
  labs(fill = "Number of Cylinders") +
  theme(legend.position = c(.25,.75)) +
  guides(fill = guide_legend(nrow = 1,
                             label.position = "bottom"
                             ))
```


```{r}
p <- ggplot(mpg, aes(x = hwy))

p + geom_density(aes(fill = factor(cyl)), alpha = .5) +
  scale_x_continuous(breaks = seq(5,40,5)) +
  scale_fill_brewer(palette = 'Reds') +
  labs(fill = "Number of Cylinders") +
  theme(legend.position = c(.25,.75)) +
  guides(fill = guide_legend(nrow = 1,
                             label.position = "bottom"
                             ))
```


### Boxplot 

```{r}
mpg |> 
  ggplot(aes(x = factor(cyl), y = hwy,
             fill = factor(cyl))) +
  geom_boxplot(show.legend = F)
```



<br>

## Q1c.
- What happens if you make a scatterplot of `class` vs. `drv`? 
  - Why is the plot not useful?


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv), alpha = .02)
```


```{r}
mpg |> 
  count(class)

mpg |> 
  count(drv)
```

### Bar charts

#### Stacked Bar Charts
```{r}
mpg |> 
  ggplot(aes(x = class, fill = drv)) +
  geom_bar()
```

#### Stacked Proportional Bar Charts
```{r}
mpg |> 
  ggplot(aes(x = class, fill = drv)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion")
```

#### Side-by-Side Bar Charts
```{r}
mpg |> 
  ggplot(aes(x = class, fill = drv)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  labs(y = "Proportion")
```


<br><br><br>

# Question 2. Aethetic Mappings


## Q2a.
- Which variables in `mpg` are categorical? 
```{r}
mpg |> 
  count(drv)
```


- Which variables are continuous?

<br>

## Q2b.
- Map a continuous variable to `color`, `size`, and `shape`. 
- How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
#| eval: false
mpg |> 
  ggplot(aes(x = displ, y = hwy,
             shape = cty)) +
  geom_point()
```


<br>

## Q2d.
What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`? 


```{r}
mpg |> 
  ggplot(aes(x = displ, y = hwy,
             color = displ < 5)) +
  geom_point(alpha = .33)
```


```{r}
mpg |> 
  mutate(cond = displ < 5) |> 
  ggplot(aes(x = displ, y = hwy,
             color = cond)) +
  geom_point(alpha = .33)
```


<br><br><br>


# Question 3. Facets

## Q3a.
What happens if you facet on a continuous variable?

```{r}
beer <- read_csv("https://bcdanl.github.io/data/beer_markets.csv")

```
```{r}
#| fig-height: 18
beer |> 
  ggplot(aes(x = log(beer_floz), y = log(price_per_floz)) ) +
  geom_point(alpha = .1) +
  geom_smooth() +
  geom_smooth(method = 'lm', color = 'red') +
  facet_grid(brand ~ .)
```

```{r}
#| eval: false
# it takes too long!
beer |> 
  ggplot(aes(x = log(beer_floz), y = log(price_per_floz)) ) +
  geom_point(alpha = .1) +
  geom_smooth() +
  geom_smooth(method = 'lm', color = 'red') +
  facet_wrap(dollar_spent ~ .)
```


<br>

## Q3b.
- What do the empty cells in the scatterplot of `displ` vs. `hwy` with `facet_grid(drv ~ cyl)` mean? 
  - How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```



<br>

## Q3c.
What plots does the following code make? What does `.` do?
```{r}
#| eval: false

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy),
             alpha = .5) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy),
             alpha = .5) +
  facet_grid(. ~ cyl)
```

<br>

## Q3d.
Consider the following faceted plot:
```{r}

ggplot(data = mpg,
       mapping = aes(x = displ, y = hwy,
                           color = class)
             ) + 
  geom_point(alpha = .5) +
  geom_smooth() +
  facet_wrap(~ class, ncol = 1) 

```

```{r}

ggplot(data = mpg,
       mapping = aes(x = displ, y = hwy,
                           color = class)
             ) + 
  geom_point(alpha = .5) +
  geom_smooth()
```

- What are the advantages to using faceting instead of the color aesthetic? 
  - What are the disadvantages? 
  - How might the balance change if you had a larger dataset?
  
  

<br>

## Q3e.
- Read `?facet_wrap`. 
  - What does `nrow` do? 
  - What does `ncol` do? 
  - What other options control the layout of the individual panels? 
  - Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments?

<br>

## Q3f.
When using `facet_grid()`, you should usually put the variable with more unique levels in the columns. Why?

<br>

## Q3g.
Use the following data.frame.

```{r}
#| echo: true
tvshows_web <- read_csv(
		'https://bcdanl.github.io/data/tvshows.csv')
```

```{r}
datatable(tvshows_web)
```

<br>

- Describe the relationship between audience size (`GRP`) and audience engagement (`PE`) using `ggplot`. 

```{r}
tvshows_web |> 
  ggplot(aes(x = GRP, y = PE)) +
  geom_point() +
  geom_smooth()
```


- Describe how the relationship between audience size (`GRP`) and audience engagement (`PE`) varies by `Genre` using `ggplot`. 

```{r}
tvshows_web |> 
  ggplot(aes(x = GRP, y = PE)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(Genre ~ .)
```


```{r}
tvshows_web |> 
  ggplot(aes(x = GRP, y = PE, color = Genre)) +
  geom_point() +
  geom_smooth(method = lm) 
```

- Explain the relationship in words. 


<br>

